---
title: "How to use xmR() and xmR_chart()"
author: "Alex Zanidean"
output: 
  html_document: 
    css: css/custom.css
    toc: yes
    toc_depth: 1
---
#Introduction

XMR control charts are useful when determining if there are significant trends in data. XMR charts have two key assumptions: one is that the measurements of value happen over time, and the other is that each measurement of time has one measurement of value. 


Take careful thought about *what* you are trying to measure with XMR. Proportions work best, headcount is okay, costs over time aren't great.

-------

#Arguments

The arguments for `xmR()` are as follows. 

- **df**: The dataframe containing the time-series data. At least 1 variable for time and one variable for measure.

- **measure**: The column containing the measure. Must be in a numeric format.

- **interval**: The interval you'd like to use to calculate the averages. Defaults to 5.

- **recalc**: Logical if you'd like it to recalculate bounds. Defaults to False for safety.


The data required for XMR charts take a specific format, with at least two columns of data - one for the time variable and another for the measurement. Like so:

```{r, message=FALSE, echo = F}
library(sRa)
set.seed(1)
Measure <- round(runif(10, min = 0.50, max = 0.75)*100, 0)
Measure <- c(Measure, round(runif(8, min = 0.75, max = 1)*100, 0))
Time <- c(2000:2017) 
example_data <- data.frame(Time, Measure)
knitr::kable(example_data, format = "markdown", align = 'c')
```

The function use on this set of data would be written like this if we wanted the boundries to recalculate. 

```{r, message=FALSE, eval = F}
xmr_data <- xmR(df = example_data, measure = "Measure", recalc = T)
```

The measure column must be written within quotes. 

Output data looks like this:
```{r, echo=F, message=FALSE}
xmr_data <- xmR(df = example_data, measure = "Measure", recalc = T) %>% select(-Order)
knitr::kable(xmr_data, format = "markdown", align = 'c')
```
-------

# Charts

That data calculation is handy, but what about visualiztion?

There is a function called `xmR_chart()` which takes that output, and generates a ggplot graphic so the user can identify trends in the data. 

The arguments for `xmR_chart()` are as follows. 

- **dataframe**: Output from xmR()

- **time**: The column containing the time variable for the x-axis.

- **measure**: The column containing the measure for the y-axis. 

- **facetvar**: Split the chart by a facet variable. Explained in the next section.


```{r, fig.height=5}
xmR_chart(dataframe = xmr_data, time = "Time", measure = "Measure")
```

-------

# Other Packages

Simple datasets like those illustrated above are common, but more common are large datasets with multiple factors. Consider the following data. How would the `xmR()` function benefit the user in this case? 

```{r, message=FALSE, echo = F}
library(sRa)
library(air)
library(tidyverse)
lers_data <- get_enrolment("FLE", c("Gender"), c("MH")) %>% 
  mutate(`Academic Year` = gsub("-20", "-", `Academic Year`))
knitr::kable(lers_data, format = "markdown", align = 'c')
```

The answer is by leveraging other R packages, namely the `tidyverse`. 

You can install and load the tidyverse by the following. 

```{r, eval = F}
install.packages("tidyverse")
library(tidyverse)
```

What this enables is a number of verb-type functions for tidying and wrangling data. This is how to use them alongside the `xmR()` function. 


-------

# Grouping and Faceting

Take our data of enrolment `lers_data` and here is how to apply the `xmR()` function to certain groups within that data. 

```{r, echo=FALSE}
lers_xmr <- lers_data %>% group_by(Gender) %>% do(xmR(., "FLE", recalc = T))
knitr::kable(lers_xmr, format = "markdown", align = 'c')
```

And as you may be able to see, the XMR data calculated on FLE **AND** by Gender, in one function instead of having to manually split the data. Now, if we tried to plot this data we would do the following. 


```{r, fig.height=7, fig.width=10}
xmR_chart(dataframe = lers_xmr, time = "Academic Year", measure = "FLE", facetvar = "Gender")
```
