---
title: "How to Use xmR() and xmR_chart() functions"
author: "Alex Zanidean"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#Introduction

XMR control charts are useful when determining if there are significant trends in data. XMR charts have two key assumptions: one is that the measurements of value happen over time, and the other is that each measurement of time has exactly one measurement of value. 


Take careful thought about *what* you are trying to measure with XMR. Proportions work best, headcount is okay, costs over time aren't great.

-------

#Arguments

The arguments for `xmR()` are as follows. 

- **df**: The dataframe containing the time-series data.

- **measure**: The column containing the measure. This must be in a numeric format.

- **interval**: The interval you'd like to use to calculate the averages. Defaults to 5.

- **recalc**: Logical if you'd like it to recalculate bounds. Defaults to False for safety.

- **reuse**: Logical: Should points be re-used in calculations? Defaults to False.


The data required for XMR charts take a specific format, with at least two columns of data - one for the time variable and another for the measurement. Like so:

```{r, message=FALSE, echo = F}
library(sRa)
library(tidyverse)
set.seed(1)
Measure <- round(runif(10, min = 0.50, max = 0.75)*100, 0)
Measure <- c(Measure, round(runif(8, min = 0.75, max = 1)*100, 0))
Time <- c(2000:2017) 
example_data <- data.frame(Time, Measure)
knitr::kable(example_data, format = "markdown", align = 'c')
```

The function use on this set of data would be written like this if we wanted the boundaries to recalculate. 

```{r, message=FALSE, eval = F}
xmr_data <- xmR(df = example_data, measure = "Measure", recalc = T)
```

The measure column must be written within quotes. 

Output data looks like this:
```{r, echo=F, message=FALSE}
xmr_data <- xmR(df = example_data, 
                measure = "Measure", 
                recalc = T) %>% 
  select(-Order)
knitr::kable(xmr_data, format = "markdown", align = 'c')
```
-------

# Charts

That data calculation is handy, the output can be saved and used in other applications. But what about visualization within R?

There is a function called `xmR_chart()` which takes that output, and generates a ggplot graphic so the user can identify trends in the data. This works well for reporting, but it also works great for quick diagnostics of your data.

The arguments for `xmR_chart()` are as follows. 

- **dataframe**: Output from xmR()

- **time**: The column containing the time variable for the x-axis.

- **measure**: The column containing the measure for the y-axis. 

- **facetvar**: Split the chart by a facet variable. Explained in the next section.


```{r, fig.height=5, fig.width=7.5}
xmR_chart(dataframe = xmr_data, time = "Time", measure = "Measure")
```

-------

# Tidyverse

Simple datasets like those illustrated above are common, but more common are large datasets with multiple factors. Consider the following data. How would the `xmR()` function benefit the user in this case? 

```{r, message=FALSE, echo = F}
library(sRa)
library(tidyverse)
`Year` <- seq(2004, 2017, 1)
Variable <- "A"
FDA <- data.frame(`Year`, Variable, check.names = F)
Variable <- "B"
FDB <- data.frame(`Year`, Variable, check.names = F)
FD <- rbind(FDA, FDB)

FD$Measure <- runif(nrow(FD))
knitr::kable(FD, format = "markdown", align = 'c')
```

The answer is by leveraging other R packages, namely the `tidyverse`. 

You can install and load the tidyverse by the following. 

```{r, eval = F}
install.packages("tidyverse")
library(tidyverse)
```

What this enables is a number of verb-type functions for tidying and wrangling data. This is how to use them alongside the `xmR()` function. 


-------

# Grouping and Faceting

Take our fake data `FD` and here is how to apply the `xmR()` function to certain groups within that data. 

```{r}
FD <- FD %>% group_by(Variable) %>% do(xmR(., "Measure", recalc = T))
```

To obtain the following:

```{r, echo=FALSE}
knitr::kable(FD, format = "markdown", align = 'c')
```

And as you may be able to see, the XMR data calculated on Measure **AND** by Variable, in one function instead of having to manually split the data. Now, if we tried to plot this data we would do the following. 


```{r, fig.height=5, fig.width=7.5, eval=F}
xmR_chart(dataframe = FD, 
          time = "Year", 
          measure = "Measure", 
          facetvar = "Variable") + 
  scale_x_discrete(breaks = scales::pretty_breaks(5))
```

Alternatively, you could use the built-in ggplot2 function for more complex faceting.

```{r, fig.height=5, fig.width=7.5}
xmR_chart(dataframe = FD, 
          time = "Year", 
          measure = "Measure") + 
  facet_wrap(~Variable) + 
  scale_x_discrete(breaks = scales::pretty_breaks(5))
```
